---
title: "Assignment 5"
subtitle: "Interactive Maps with Leaflet"
output: html_document
author: "Kaarunya Mohanathas"
date: "2023-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
The object of this assignment is to create an interactive map using the <a href="https://github.com/Digital-Methods-HASS/MapsWithLeaflet">Leaflet</a> tutorial and following the steps as stated in the StartWithLeaflet.R file from the <a href="https://github.com/Digital-Methods-HASS/MapsWithLeaflet/tree/main">MapsWithLeaflet repository</a>.

**Activating packages**
```{r}
library(leaflet)
library(htmlwidgets)
library(tidyverse)
library(googlesheets4)
library(leaflet)
```
### Task 1
**Create a Danish equivalent of AUSmap with esri layers, but call it DANmap**</br>
I used the StartWithLeaflet.R file as a template, and created an object named **l_dan** and assigned it the proper coordinates. Afterwards, an object called **esri** is created for selecting background layers for the mini-map. The provider tiles are chosen from the <a href="https://leaflet-extras.github.io/leaflet-providers/preview/">Leaflet Provider Demo</a> site.

```{r task_1_1}
l_dan <- leaflet() %>%
  setView(11.2149938, 56.1510991, zoom = 6)

esri <- grep("^Esri", providers, value = TRUE)

for (provider in esri) {
  l_dan <- l_dan %>% addProviderTiles("Esri.WorldStreetMap")
}
```

By using the code from **AUSmap** with *slight* alterations, **DANmap** is created.

```{r task_1_2}
DANmap <- l_dan %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = TRUE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }")

DANmap
```

### Task 2
**Read in the googlesheet data you and your colleagues populated with data into the DANmap object you created in Task 1.**</br>
To begin with, the data from the CapturingCoordinates googlesheet is read into an object called **places**. 
```{r task_2_1}
gs4_deauth()

places <- read_sheet("https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918",
                     col_types = "cccnncnc", range = "DigitalMethods")

glimpse(places)
```
The data is loaded in the map to check if everything looks as it is supposed to.
```{r}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description)
```

Then, the **DANmap** object from Task 1 is populated with the data, using the code from StartWithLeaflet.R.
```{r task_2_2}
DANmap %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Placename)
```

*Note: I have added the Placename data from the spreadsheet as popups here instead of the Description variable.*

### Task 3
**Can you cluster the points in Leaflet? Google "clustering options in Leaflet"**</br>
By using the clusterOptions functions, the markers are clustered so as to appear a bit more visually appealing.
```{r}
DANmap %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Placename,
             clusterOptions = markerClusterOptions())
```

Source: <http://rstudio.github.io/leaflet/markers.html>

### Task 4
**What is the map good for and what is it not?**</br>
The map above is a good overview of Danish attractions as noted by the students in our class. You can use it to discern where they are located and approx. how far the attractions are from one another.</br>
However, it is not an exhaustive list and many notable attraction may not appear. It is not a very detailed map either, and the different types of attractions cannot be differentiated from one another.</br>

### Task 5
**Find out how to display notes and classifications in the map.**</br>
Assuming here, that notes and classifications can be in the form of popups, we can use the inbuilt popup function in Leaflet.
When the markers in the map above are clicked, popups with the names of the attractions will appear. This is due to **popup = places$Placename**.</br>
The **addPopups** function can be used to add standalone popups.</br>
Source: <https://rstudio.github.io/leaflet/popups.html>